---
title: "Task 1 - Sierra Amphibians"
author: "Alex Ehrens"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# attach packages
library(tidyverse)
library(janitor)
library(patchwork)
library(readxl)
library(lubridate)
```

#### Step 1 - Read in Sierra amphibians Excel spreadsheet
```{r}
# read in file and change column headers to lower snake case
amphibians <- read_excel("sierra_amphibians.xlsx") %>% 
  clean_names()
```

#### Step 2 - Wrangle the data to get a total mountain yellow-legged frog (*Rana muscosa*) count each year
```{r}
# filter to include only Rana muscosa (RAMU) species and exclude the "EggMass" life stage, then convert survey_date to date format and pull out just the year
amphibians_ramu <- amphibians %>% 
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage != "EggMass") %>%
  mutate(survey_date = ymd(survey_date)) %>% 
  mutate(year = year(survey_date))

# group by life stage and year, then get a count of Rana muscosa
amphibians_ramu_count <- amphibians_ramu %>% 
  count(year, amphibian_life_stage, wt = amphibian_number) %>% 
  rename(count = n)
```

#### Step 3 - Produce plot of mountain yellow-legged frog (*Rana muscosa*) count
```{r}
count_plot1 <- ggplot(data = amphibians_ramu_count,
                      aes(x = year, y = count, color = amphibian_life_stage, fill = amphibian_life_stage)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  theme_bw() +
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(labels = c(1996, 1998, 2000, 2002), breaks = c(1996, 1998, 2000, 2002)) +
  facet_wrap(~amphibian_life_stage, scales = "free") +
  labs(title = expression(paste("Counts of mountain yellow-legged frogs (", italic("Rana muscosa"), ") by year and life stage")),
       fill = "Life Stage",
       color = "Life Stage",
       x = "Year of Survey",
       y = "Individuals Observed")

count_plot1
```

#### Step 4 - Wrangle data to get total counts of adult and subadult *Rana muscosa* in the 5 lakes with greatest total observed counts
```{r}
# filter original amphibians dataset to only include adult and subadult frogs, then reformat lake ID
amphibians_lakes <- amphibians %>% 
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage == c("Adult", "SubAdult")) %>% 
  mutate(lake_id = paste0("Lake ", lake_id))

# group by lakes and summarize to get counts of frogs by lake, then 
amphibians_top5_lakes <- amphibians_lakes %>% 
  count(lake_id, wt = amphibian_number) %>% 
  rename(count = n) %>% 
  slice_max(count, n = 5)
```

#### Step 5 - Produce plot of lakes with top 5 most observed counts of adult and subadult *Rana muscosa*
```{r}
count_plot2 <- ggplot(data = amphibians_top5_lakes,
                      aes(x = fct_reorder(lake_id, -count), y = count)) +
  geom_col(aes(color = fct_reorder(lake_id, count), fill = fct_reorder(lake_id, count)),
           show.legend = FALSE) +
  scale_fill_brewer(palette = "YlGnBu") +
  scale_color_brewer(palette = "YlGnBu") +
  theme_minimal() +
  labs(title = expression(paste("Lakes with the highest observed counts of adult and subadult ", italic("Rana muscosa"))),
       x = "Lake Identification Number",
       y = "Individuals Observed")

count_plot2
```

#### Step 6 - use `{patchwork}` package to combine two graphs into single figure 
```{r}
patchwork_plot <- count_plot1 / count_plot2 & 
  plot_annotation(caption = "Plots of total observed counts of mountain yellow-legged frogs (Rana muscosa) in the lakes and ponds of the Sierra Nevada \nmountains of California. Plot 1 (top) displays counts of individuals by year of survey and amphibian life stage (Adult in red, \nSubAdult in blue, and Tadpole in green). Plot 2 (bottom) shows the total observed counts of combined Adults and SubAdults \nin the 5 lakes with the highest number of observed individuals. Data from The Sierra Lakes Inventory Project, 2020.") &
  theme(plot.caption = element_text(hjust = 0))

patchwork_plot
```

##### End Task 1